---
title: "Poisson / Negative Binomial Modeling"
author: "MSup"
output: html_document
---
##### Code executed: `r Sys.Date()`

```{r, echo=FALSE}
load("../data/dfTrain.Rda") 
load("../data/dfTest.Rda") 
library(ggplot2)
```

### Outcome: Count of Bike Rentals 
```{r, echo=FALSE, comment="", fig.height=5, fig.width=8}
summary(dfTrain$count)
(ggplot(dfTrain, aes(x=count)) + geom_histogram(binwidth=2, colour="darkblue", fill="white") + ggtitle("Distribution of Count") + theme_bw() + geom_vline(aes(xintercept=median(count, na.rm=T)), color="sienna1", linetype="dashed", size=1))
```

***

### Saturated Model: Poisson 
#### All predictors with bivariate p-val < 0.20 are included in the model 
```{r, echo=FALSE, comment=""}
pois <- glm(count ~ season_char + month + day 
            + hour + am_pm + as.factor(weather)
            + atemp + humidity + windspeed, 
            family = poisson, data = dfTrain)
summary(pois)
# getting rid of scientific notation
options(scipen=999)
```
Checking the overdispersion, the residual deviance should be equal to the residual degrees of freedom if the Poisson errors assumption is met:
Residual deviance = `r pois$deviance`  
Residual df = `r pois$df.residual`  
Model is overdispersed (a lot!!!), will use Negative Binomial model to adjust for it.  

***

### Saturated Model: Negative Binomial 
#### All predictors with bivariate p-val < 0.20 are included in the model 
```{r, echo=FALSE, comment=""}
# returning scientific notation
options(scipen=0)
library(MASS)
nb <- glm.nb(count ~ season_char + month + day 
            + hour + am_pm + as.factor(weather)
            + atemp + humidity + windspeed, 
            data = dfTrain)
summary(nb)
```







