---
title: "Bike Demand Prediction Notes"
output: html_document
---
Problem Statement
-----------------
You are provided hourly rental data spanning two years. For this competition, the training set is comprised of the first 19 days of each month, while the test set is the 20th to the end of the month. You must predict the total count of bikes rented during each hour covered by the test set, using only information available prior to the rental period.

Data Fields
-----------
+ **datetime**
    hourly date + timestamp  
+ **season**
    * 1 = spring 
    * 2 = summer 
    * 3 = fall
    * 4 = winter 
+ **holiday** 
  whether the day is considered a holiday
+ **workingday** whether the day is neither a weekend nor holiday
+ **weather**  
    * 1: Clear, Few clouds, Partly cloudy, Partly cloudy
    * 2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist
    * 3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds
    * 4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog 
+ **temp** - temperature in Celsius
+ **atemp** - "feels like" temperature in Celsius
+ **humidity** - relative humidity
+ **windspeed** - wind speed
+ **casual** - number of non-registered user rentals initiated
+ **registered** - number of registered user rentals initiated
+ **count** - number of total rentals

Workspace 
----------

Loading the training data set
```{r}
data.train = read.csv('data\\train.csv')
summary(data.train)

data.test = read.csv('data\\test.csv')
summary(data.test)
```

Generate prediction file function. Use this method to generate file.
For example `generateKaggleOutput('test.csv', testData, predictionData)'
```{r}
generateKaggleOutput = function(fileName="output.csv", test, prediction){
  pred.output = cbind(datetime=format(as.POSIXlt(test[,1])),count=prediction)
  write.csv(pred.output, file=paste("submissions\\",fileName, sep=""), row.names=FALSE)
}
pkgTest = function(x) {
  if(x %in% rownames(installed.packages()) == FALSE) {
    install.packages(x)
  }
}
```


Base Line data
```{r}
baseCount = as.integer(mean(data.train[,'count']))
generateKaggleOutput("test.csv", data.test, baseCount)

```

Check for the baseline RMSLE
```{r}
pkgTest("Metrics")
library("Metrics")
rmsle(baseCount, data.train$count)
```



```{r}
par(mfrow = c(3,3))
season = data.train[,2]
hist(season) # season categorical data
table(season)
plot(season)
boxplot(season)
holiday = data.train[,3] # binary data
hist(holiday) # Holiday
table(holiday)
plot(holiday)
boxplot(holiday)
workingday = data.train[,4] # binary data
hist(workingday) # workingday 
table(workingday)
plot(workingday)
boxplot(workingday)
par(mfrow = c(3,3))
weather = data.train[,5] # categorical data
hist(weather) # weather
table(weather)
plot(weather)
boxplot(weather)
temp = data.train[,6] # continuous data or quantitative data
hist(temp) # temp
table(temp)
plot(temp)
boxplot(temp)
atemp = data.train[,7] # continous data or quantitative data
hist(atemp) # atemp "feels like"
table(atemp)
plot(atemp)
boxplot(atemp)
par(mfrow = c(3,3))
humidity = data.train[,8] # humidity continuous data or quantitative data
hist(humidity)
table(humidity)
plot(humidity)
boxplot(humidity)
windspeed = data.train[,9] # windspeed continuous data or quantitative data
hist(windspeed)
table(windspeed)
plot(windspeed)
boxplot(windspeed)
casual = data.train[,10] # casual quantitative data looks like long tail
hist(casual)
plot(casual)
boxplot(casual)
par(mfrow = c(2,3))
registered = data.train[,11] # registered quantitative data long tail
hist(registered)
plot(registered)
boxplot(registered)
count = data.train[,12] # count quantitative data 
hist(count)
plot(count)
boxplot(count)
```

Trying something new here
```{r}
load("dfTrain.Rda")
summary(dfTrain)


```

