---
title: "Bivariate Models"
output: pdf_document
---

#### Outcome: Count  

>Code executed: `r Sys.Date()`

```{r, echo=FALSE, tidy=TRUE, comment=""}
#loading data 
load("./data/dfTrain.Rda") 
library(ggplot2)
# function for pretty p-value output 
prettyPVal <- function(pval) {
  if (is.na(pval)) {
    "NA"
  } else {
    log10pval <- log10(pval)
    if (log10pval < -5) {
      sprintf("< 0.00001")
    } else if (log10pval < -4) {
      sprintf("< 0.0001")
    } else if (log10pval < -3) {
      sprintf("< 0.001")
    } else {
      sprintf("%0.3f", round(pval, digits=3))
    }
  }
}
```

***

<center> <h2>  Seasons vs. Count </h2> </center>
```{r, echo=FALSE, tidy=TRUE, comment="", fig.height=4, fig.width=7}
(ggplot(dfTrain, aes(x=season_char, y=count, fill=season_char)) + geom_boxplot()
 + ggtitle("Seasons vs. Count"))
fit <- lm(count ~ season_char, data=dfTrain)
summary(fit)
pval <- (anova(fit)$'Pr(>F)'[1])
```
#### Global p-val `r prettyPVal(pval)`

***

<center> <h2> Months vs. Count </h2> </center>
```{r, echo=FALSE, tidy=TRUE, comment="", fig.height=4, fig.width=7}
(ggplot(dfTrain, aes(x=as.factor(month), y=count, fill=as.factor(month))) 
 + geom_boxplot() + ggtitle("Month vs. Count"))
fit <- lm(count ~ month, data=dfTrain)
summary(fit)
pval <- (anova(fit)$'Pr(>F)'[1])
```
#### Global p-val `r prettyPVal(pval)`

***

<center> <h2> Holiday vs. Count </h2> </center>
```{r, echo=FALSE, tidy=TRUE, comment="", fig.height=4, fig.width=7}
(ggplot(dfTrain, aes(x=as.factor(holiday), y=count, fill=as.factor(holiday))) 
 + geom_boxplot() + ggtitle("Holiday vs. Count"))
fit <- lm(count ~ holiday, data=dfTrain)
summary(fit)
pval <- (anova(fit)$'Pr(>F)'[1])
```
#### Global p-val `r prettyPVal(pval)`

***

<center> <h2> Workingday vs. Count </h2> </center>
```{r, echo=FALSE, tidy=TRUE, comment="", fig.height=4, fig.width=7}
(ggplot(dfTrain, aes(x=as.factor(workingday), y=count, fill=as.factor(workingday))) 
 + geom_boxplot() + ggtitle("Workingday vs. Count"))
fit <- lm(count ~ workingday, data=dfTrain)
summary(fit)
pval <- (anova(fit)$'Pr(>F)'[1])
```
#### Global p-val `r prettyPVal(pval)`

***

<center> <h2> Week Day vs. Count </h2> </center>
```{r, echo=FALSE, tidy=TRUE, comment="", fig.height=4, fig.width=7}
(ggplot(dfTrain, aes(x=as.factor(week_day), y=count, fill=as.factor(week_day))) 
 + geom_boxplot() + ggtitle("Week_Day vs. Count") +theme(axis.title.x = element_blank())
 + scale_x_discrete(limits=c("Monday", "Tuesday", "Wednesday", 
                             "Thursday", "Friday", "Saturday", "Sunday")))
fit <- lm(count ~ week_day, data=dfTrain)
summary(fit)
pval <- (anova(fit)$'Pr(>F)'[1])
```
#### Global p-val `r prettyPVal(pval)`

***

<center> <h2> Hour vs. Count </h2> </center>
```{r, echo=FALSE, tidy=TRUE, comment="", fig.height=4, fig.width=7}
(ggplot(dfTrain, aes(x=hour, y=count)) + geom_point(shape=1) 
 + geom_smooth(method=lm, se=FALSE) + ggtitle("Hour vs. Count"))
fit <- lm(count ~ hour, data=dfTrain)
summary(fit)
pval <- (anova(fit)$'Pr(>F)'[1])
```
#### Global p-val `r prettyPVal(pval)`

***

<center> <h2> AM/PM vs. Count </h2> </center>
```{r, echo=FALSE, tidy=TRUE, comment="", fig.height=4, fig.width=7}
(ggplot(dfTrain, aes(x=as.factor(am_pm), y=count, fill=as.factor(am_pm))) 
 + geom_boxplot() + ggtitle("AM/PM vs. Count"))
fit <- lm(count ~ am_pm, data=dfTrain)
summary(fit)
pval <- (anova(fit)$'Pr(>F)'[1])
```
#### Global p-val `r prettyPVal(pval)`

***

<center> <h2> Weather vs. Count </h2> </center>
#### Wether Types:  
##### 1. Clear, Few clouds, Partly cloudy, Partly cloudy   
##### 2. Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist   
##### 3. Light Snow, Light Rain + Thunderstorm + Scatt clouds, Light Rain + Scatt clouds   
##### 4. Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog   
```{r, echo=FALSE, tidy=TRUE, comment="", fig.height=4, fig.width=7}
(ggplot(dfTrain, aes(x=as.factor(weather), y=count, fill=as.factor(weather))) 
 + geom_boxplot() + ggtitle("Weather vs. Count"))
fit <- lm(count ~ as.factor(weather), data=dfTrain)
summary(fit)
pval <- (anova(fit)$'Pr(>F)'[1])
```
#### Global p-val `r prettyPVal(pval)`

***

<center> <h2> Temperature vs. Count </h2> </center>
```{r, echo=FALSE, tidy=TRUE, comment="", fig.height=4, fig.width=7}
(ggplot(dfTrain, aes(x=temp, y=count)) + geom_point(shape=1) 
 + geom_smooth(method=lm, se=FALSE) + ggtitle("Temp vs. Count"))
fit <- lm(count ~ temp, data=dfTrain)
summary(fit)
pval <- (anova(fit)$'Pr(>F)'[1])
```
#### Global p-val `r prettyPVal(pval)`

***

<center> <h2> Ambient Temperature vs. Count </h2> </center>
```{r, echo=FALSE, tidy=TRUE, comment="", fig.height=4, fig.width=7}
(ggplot(dfTrain, aes(x=atemp, y=count)) + geom_point(shape=1) 
 + geom_smooth(method=lm, se=FALSE) + ggtitle("Ambient Temp vs. Count"))
fit <- lm(count ~ atemp, data=dfTrain)
summary(fit)
pval <- (anova(fit)$'Pr(>F)'[1])
```
#### Global p-val `r prettyPVal(pval)`

***

<center> <h2> Humidity vs. Count </h2> </center>
```{r, echo=FALSE, tidy=TRUE, comment="", fig.height=4, fig.width=7}
(ggplot(dfTrain, aes(x=humidity, y=count)) + geom_point(shape=1) 
 + geom_smooth(method=lm, se=FALSE) + ggtitle("Humidity vs. Count"))
fit <- lm(count ~ humidity, data=dfTrain)
summary(fit)
pval <- (anova(fit)$'Pr(>F)'[1])
```
#### Global p-val `r prettyPVal(pval)`

***

<center> <h2> Windspeed vs. Count </h2> </center>
```{r, echo=FALSE, tidy=TRUE, comment="", fig.height=4, fig.width=7}
(ggplot(dfTrain, aes(x=windspeed, y=count)) + geom_point(shape=1) 
 + geom_smooth(method=lm, se=FALSE) + ggtitle("Windspeed vs. Count"))
fit <- lm(count ~ windspeed, data=dfTrain)
summary(fit)
pval <- (anova(fit)$'Pr(>F)'[1])
```
#### Global p-val `r prettyPVal(pval)`